Appendix 1: Preregistered analysis with simulated data

### 2025-04-08 preregistration.Rmd
sampling

2. Psychological ownership over FRCBS increases with donation experience. That is the lower bound of effect estimate of 95% standardized confidence or credible interval of direct effect of past donation behavior to psychological ownership is higher than zero.

@@

Cross sectional survey stud ymodelled with a structural equation model.

@@

A set of blood donors will be randomly selected from all active blood donors in FRBS register.

@@

An e-mail inviting to fill an online questionnaire will be sent to the randomly selected blood donors. No recompensation will be offered to participants.Participants must be between 18 and 65 years old.

@@

Our target sample size is 1000 participants.

@@

Enter your response here.

@@

All likert scale measured variables will be treated as numeric as well number of life time donations (bd02) and number of donations in last years (bd03). Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".

For sum score model latent variables will be formed by summing measured items

For structural equation models using latent variables

@@

Enter your response here.

Items
<table>
<tr>
<td></td>
<td>Psychological ownership</td>
</tr>
<tr>
<td>
po01 1. I feel a very high degree of personal ownership of donating blood.
</td>
</tr>
<td>
po02 2. I feel like donating blood belongs to me.
</td> </tr>
<tr>
<td>
po03 3. I feel like I own donating blood.
</td> </tr>
<tr> <td></td>

<td>
Self-identity
</td> </tr>
<tr>
<td>
si01 1. Donating blood is important to me.
</td> <tr>
si02 2. I am like the kind of person who donates blood.
</td> </tr>
<tr> 
<td>
si03 3. Donating blood is an important part of who I am.
</td> </tr>
<tr> <td></td>

<td>
Intention
</td></tr>
<tr>
<td>
it01 1. I would like to donate blood in the future.
</td>   </tr>
<tr>
<td>
it02 2. I intend to donate blood in the future.
</td> </tr>
<tr>
<td>
it03 3. I intend to make more than just a one-off blood donation.
</td></tr>
<tr><td></td>
<td>
Intention to donate for emergency
</td></tr>
<tr>
<td>
ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service.
</td></tr>
<tr><td></td> 
<td>
Antecedents of psychological ownership
</td></tr>
<tr><td></td>

<td>
Control
</td></tr>
<tr>
<td>
ac01  1. I have influence over the things that affect me while donating blood.
</td></tr>
<tr>
<td>
ac02 2. I control the location and scheduling for donating blood.
</td></tr>
<tr>
<td>
ac03 3. I influence decisions related to my blood donation.
</td></tr>
<tr>
<td>
ac04 4. In general, I have control over donating blood.
</td></tr>
<tr><td></td>
<td>
Intimate knowledge
</td></tr>
<tr>
<td>
ai01 1. I am intimately familiar with what is going on with regard to donating blood.
</td></tr>
<tr>
<td>
ai02 2. I have a depth of knowledge as it relates to donating blood.
</td></tr>
<tr>
<td>
ai03 3. I have a comprehensive understanding of donating blood.
</td></tr>
<tr>
<td>
ai04 4. I have a broad understanding of donating blood.
</td></tr>
<tr><td></td>

<td>
Self-Investment
</td></tr>
<td>
iv01 1.	I have invested a major part of myself into donating blood.
</td></tr>
<tr>
<td>
iv02 2. I have invested a significant amount of my time into donating blood.
</td></tr>
<tr>
<td>
iv03 3. Overall, I invested a lot into donating blood.
</td></tr>
<tr><td></td>

@@

sum scores and latent variables

@@

SEM include R from preregistartion code

@@

as numeric

@@

Among

@@

All donors answering to question de01 other than "man" or "women" will be excluded as donation policies depend on sex. All donors answering to question bd01 "no" will be excluded as we are interested in person eligible to donate.

>>>>

2. Psychological ownership over FRCBS increases with donation experience. That is the lower bound of effect estimate of 95% standardized confidence or credible interval of direct effect of past donation behavior to psychological ownership (a1) is higher than zero.

@@

Cross sectional survey study modelled with a structural equation model.

@@

A set of blood donors will be randomly selected from all blood donors in FRCBS register who have donated at least once during the 730 preceding the day the questionnaire e-mail will be sent i.e. 5.5.2025 .

@@

An e-mail inviting to fill an online questionnaire will be sent to the randomly selected blood donors. No recompensation will be offered to participants.Participants must be between 18 and 72 years old.

@@

Our target sample size is 1000 participants. Based on previous participation rates an email will be sent to 5000 random donors.

@@

The online questionnaire will be open for 2 weeks i.e. from 5.5.2025 - to 18.5.2025.

@@

@@

All Likert items are scored 1 (strongly disagree) to 7 (strongly agree).

Items:
<table>
<tr>
<td></td>
<td>Psychological ownership</td>
</tr>
<tr>
<td>
po01 1. I feel a very high degree of personal ownership of donating blood. [Likert]
</td>
</tr>
<td>
po02 2. I feel like donating blood belongs to me. [Likert]
</td> </tr>
<tr>
<td>
po03 3. I feel like I own donating blood. [Likert]
</td> </tr>
<tr> <td></td>

<td>
Self-identity
</td> </tr>
<tr>
<td>
si01 1. Donating blood is important to me. [Likert]
</td> <tr>
si02 2. I am like the kind of person who donates blood. [Likert]
</td> </tr>
<tr> 
<td>
si03 3. Donating blood is an important part of who I am. [Likert]
</td> </tr>
<tr> <td></td>

<td>
Intention
</td></tr>
<tr>
<td>
it01 1. I would like to donate blood in the future. [Likert]
</td>   </tr>
<tr>
<td>
it02 2. I intend to donate blood in the future. [Likert]
</td> </tr>
<tr>
<td>
it03 3. I intend to make more than just a one-off blood donation. [Likert]
</td></tr>
<tr><td></td>
<td>
Intention to donate for emergency
</td></tr>
<tr>
<td>
ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service. [Likert]
</td></tr>
<tr><td></td> 
<td>
Antecedents of psychological ownership
</td></tr>
<tr><td></td>

<td>
Control
</td></tr>
<tr>
<td>
ac01  1. I have influence over the things that affect me while donating blood. [Likert]
</td></tr>
<tr>
<td>
ac02 2. I control the location and scheduling for donating blood. [Likert]
</td></tr>
<tr>
<td>
ac03 3. I influence decisions related to my blood donation. [Likert]
</td></tr>
<tr>
<td>
ac04 4. In general, I have control over donating blood. [Likert]
</td></tr>
<tr><td></td>
<td>
Intimate knowledge
</td></tr>
<tr>
<td>
ai01 1. I am intimately familiar with what is going on with regard to donating blood. [Likert]
</td></tr>
<tr>
<td>
ai02 2. I have a depth of knowledge as it relates to donating blood. [Likert]
</td></tr>
<tr>
<td>
ai03 3. I have a comprehensive understanding of donating blood. [Likert]
</td></tr>
<tr>
<td>
ai04 4. I have a broad understanding of donating blood. [Likert]
</td></tr>
<tr><td></td>

<td>
Self-Investment
</td></tr>
<td>
iv01 1.	I have invested a major part of myself into donating blood. [Likert]
</td></tr>
<tr>
<td>
iv02 2. I have invested a significant amount of my time into donating blood. [Likert]
</td></tr>
<tr>
<td>
iv03 3. Overall, I invested a lot into donating blood. [Likert]
</td></tr>
<tr><td></td>

@@

For sum score model latent variables will be formed by summing measured items:

For structural equation models using latent variables the same variables will be used as constituent items of the latent variable. Instead of using just question bd02 as in the original publication a donation history latent variable will be formed as:

Blood donation history = bd02 + bd03 + bd04

@@

We will model the data with a structural equation model. The exact model, R code to run it and example analysis with simulated data can be found in Appendix 1.

@@

Number of life time donations (bd02) and number of donations in last years (bd03) and likert scale measured variables (po01-3, si01-03, it01-03, ac01-04, iv01-03) will be treated as numeric. Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".

@@

In supplement we will report results from all models, but as the main results we will report the model with highest Comparative Fit Index (CFI). 95 % confidence or credible intervals will be used to decide if an effect is considered significant or not.

Further details of the models can be found in Appendix 1.

@@

All donors answering to question de01 other than "man" or "women" will be excluded as donation policies depend on sex. All donors answering to question bd01 "no" will be excluded as we are interested in persons eligible to donate.

### 2025-04-08 preregistration_code.Rmd
sampling

data.forest  <- fit.po03ss.best %>% 
  mutate(estmet = "Sum score\nbootstrapped") %>%  
  filter(op == "~"| op == ":=") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(estmet = "Sum score") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
    estmet = as.factor(estmet),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
                             levels=levels(estmet))
  )

>>>>

data.forest  <- fit.po03ss.best %>% 
  mutate(estmet = "Sum score\nbootstrapped") %>%  
  filter(op == "~"| op == ":=") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(estmet = "Sum score") %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
    term =str_replace_all(term,"bd02","blooddonor:"),
    term =str_replace_all(term,"blooddonor:","blooddonor"),
    estmet = as.factor(estmet),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
                             levels=levels(estmet))
  )
